NumPy---Array Computation for Python

### 2019-12-18 authors.tex
Add abstract
ENH: Add initial draft of random
Add random section draft

\author[1]{Jane Doe}
\affil[1]{Berkeley Institute for Data Science, University of California, Berkeley, CA 94720, USA}

>>>>

\author[1]{Charles R Harris}
\author[2]{Travis E. Oliphant}
\author[3,*]{St\'efan J. van der Walt}
\author[3]{Warren Weckesser}
\author[3,4,*]{K. Jarrod Millman}
\author[5]{Kevin Sheppard}
\affil[1]{Independent Researcher}
\affil[2]{Quansight LLC, Austin, TX, USA}
\affil[3]{Berkeley Institute for Data Science, University of California, Berkeley, CA 94720, USA}
\affil[4]{Division of Biostatistics, University of California, Berkeley, CA 94720, USA}
\affil[*]{numpy.articles@gmail.com}
\affil[5]{kevin.sheppard@economics.ox.ac.uk, Department of Economics, University of Oxford, Oxford, OX1 3UQ, UK}

### 2019-12-18 paper.tex
ENH: Add initial draft of random
Add random section draft

>>>>

The NumPy random module provides pseudorandom numbers from a wide range of
distributions. In legacy versions of NumPy, simulated random values are produced
by a RandomState object that: handles seeding and state initialization; wraps
the core pseudorandom number generator based on a 32-bit implementation of
MT19937; interfaces with the underlying code that transforms random bits into
deviates from other distributions; and supplies a singleton instance exposed in
the root of the random module.

The RandomState object makes a compatibility guarantee so that a fixed seed and
sequence of function calls produce the same set of values. This guarantee has
slowed progress since improving the underlying code requires extending the API
with additional keyword arguments. This guarantee continues to apply to
RandomState.

NumPy 1.17 introduced a new API for generating random numbers that use a more
flexible structure that can be extended by libraries or end-users. The new API
is build using components separate the tasks needed when generating random
variates. Pseudorandom bits are generated by a bit generator. These bits are
then transformed into variates from complex distributions by a generator.
Finally, seeding is handled by an object that produces sequences of high-quality
initial values.

Bit generators are simple classes that manage the state of an underlying
pseudorandom number generator. NumPy ships with four bit generators. The default
bit generator is a 64-bit implementation of the Permuted Congruential Generator
\cite{pcg64} (\code{PCG64}). The three other bit generators are a 64-bit version
of the Philox generator\cite{random123} (\code{Philox}), Chris Doty-Humphrey's
Small Fast Chaotic generator\cite{practrand} (\code{SFC64}), and the 32-bit
Mersenne Twister\cite{mt19937} (\code{MT19937}) which has been used in older
versions of NumPy.\footnote{The randomgen project supplies a wide range of
alternative bit generators such as a cryptographic counter-based generators
(\code{AESCtr}) and generators that expose hardware random number generators
(\code{RDRAND})\cite{randomgen}.} Bit generators export a capsule containing
pointers to functions that produce 64 bits, 32 bits, or a random double. They
also expose CFFI and ctypes interfaces to the same three functions.

The Generator consumes one of the bit generators and produces variates from
complicated distributions. Many improved methods for generating random variates
from common distributions were implemented, including the Ziggurat method for
normal, exponential and gamma variates\cite{ziggurat}, and Lemire's method for
bounded random integer generation\cite{lemire}. The Generator is more similar to
the legacy RandomState, and its API is substantially the same. The key
differences all relate to state management, which has been delegated to the bit
generator. The Generator does not make the same stream guarantee as the
RandomState object, and so variates may differ across versions as improved
generation algorithms are introduced.\footnote{Despite the removal of the
compatibility guarantee, simple reproducibility across versions is encouraged,
and minor changes that do not produce meaningful performance gains or fix
underlying bug are not generally adopted.}

Finally, a SeedSequence is used to initialize a bit generator. The seed sequence
can be initialized with no arguments, in which case it reads entropy from a
system-dependent provider, or with a user-provided seed. The seed sequence then
transforms the initial set of entropy into a sequence of high-quality
pseudorandom integers, which can be used to initialize multiple bit generators
deterministically. A key design goal of a seed sequence was to be splittable in
the sense that a seed sequence can be used to produce child sequences that are
distinct from their parents or other ancestors. This capability allows a seed
sequence to be used in large distributed applications where the number of
workers required is not known. The sequences generated from the same initial
entropy and the same splits are fully deterministic to ensure reproducibility.

The three components are combined to construct a complete random number
generator.

\begin{lstlisting}
from numpy.random import Generator, PCG64, SeedSequence
seq = SeedSequence(10304245474441179933310169597522947680)
pcg = PCG64(seq)
gen = Generator(pcg)
\end{lstlisting}

\noindent This approach retains access to the seed sequence which can then be used to
spawn additional generators.

\begin{lstlisting}
children = seq.spawn(2)
gen_0 = Generator(PCG64(children[0]))
gen_1 = Generator(PCG64(children[1]))
\end{lstlisting}

\noindent While this approach retains complete flexibility, the method
\code{np.random.default\_rng} can be used to instantize a Generator when
reproducibility is not needed.

The final goal of the new API is to improve extensibility. RandomState is a
monolithic object that obscures all of the underlying state and functions. The
component architecture is one part of the extensibility improvements. The
underlying functions (written in C) which transform the output of a bit
generator to other distributions are available for use in CFFI. This allows the
same code to be run in both NumPy and dependent that can consume CFFI, e.g.,
numba. Both the bit generators and the low-level functions can also be used in C
or Cython code.\footnote{As of 1.18.0, this scenario requires access to the
NumPy source. Alternative approaches that avoid this extra step are being
explored.}
